---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# LightFitR

<!-- badges: start -->
<!-- badges: end -->

The goal of LightFitR is to allow scientists to program complex light regimes with confidence.

## Installation

You can install the development version of LightFitR from [GitHub](https://github.com/ginavong/LightFitR) with:

``` r
# install.packages("devtools")
devtools::install_github("ginavong/LightFitR")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(LightFitR)

# Prep variables
calib <- LightFitR::calibration
times <- LightFitR::time_vector
target_irradiance <- LightFitR::target_irradiance

# Run function
makeRegime(times, target_irradiance, calib$led, calib$wavelength, calib$intensity, calib$irradiance)
```


# Devnotes

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this.

You can also embed plots, for example:

```{r pressure, echo = FALSE}
#plot(pressure)
```

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.
