---
title: "LightFitR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{LightFitR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(LightFitR)
```

# Introduction

LightFitR is an R package for designing complex light regimes with LED lights. 
Often, these light fixtures are programmed with 'intensity' units, which often does not scale linearly with the actual measured light output from the light fixtures.
Further, if using multiple wavelength channels, there will often be bleedthrough between the channels, affecting the quality and quantity of light received by your experimental subjects.
Our package aims to combat both of these challenges.
It takes calibration data and user-defined target irradiances and it tells  you what intensities to use in order to achieve those irradiances.

**Note that this package does not support broad spectrum white LEDs**

## Terminology

This packages uses 'intensity' to mean the unitless settings that the light fixture uses.
'irradiance' means the measured light output from the fixture.This can be in the users' preferred light units, provided it is used consistently throughout.
'regime' refers to a repeating daily schedule that the light fixtures run on.
'event' refers to the point in the light regime when the lights change intensity.

# Getting started

You will need several inputs: 

- calibration data from your light fixture
- target irradiances that you want your lights to achieve.
- timepoints for your events

## Calibration data

Your calibration data must have these 4 columns:

- LED channel: which channel the given measurement corresponds to
- Intensity: the intensity setting used on that channel to achieve the measurement
- Wavelength: the wavelength of the measurement (if using a light meter which doesn't measure irradiance across multiple wavelengths, set the wavelength to the peak wavelength of the channel)
- Irradiance: The measured light output on that LED channel at that intensity. This can be in any units your measurement device uses, provided you use the same units throughout.

For more information about how to collect calibration data, see https://doi.org/10.1101/2025.06.06.658293

Here is an example of what the calibration data could look like:
```{r}
calibration <- LightFitR::calibration
head(calibration)
```

## Target irradiance

The target irradiances are a set of irradiances that you wish to achieve for your experiment.

This should be a matrix, with rows representing LED channels and columns representing timepoints or events.

For example:

```{r}
target <- LightFitR::target_irradiance
print(target)
```

## Event timepoints

This should be a vector with a length the same as the column number of your target irradiances.

Each event timepoint should be in the POSIXct format. 
The date can be arbitrary as the package only cares about the timestamp.
We recommend the `lubridate` package for working with times.

For example:
```{r}
times <- LightFitR::time_vector
print(times)
```

## Creating the regime of intensities

Now you have all the inputs, let's make a regime:

```{r}
regime <- makeRegime(times, target, calibration$led, calibration$wavelength, calibration$intensity, calibration$irradiance)
print(regime)
```

As you can see, the `makeRegime` function automatically creates a heatmap of your regime to allow the user to sanity check the intensity output.

It also creates a matrix of intensities for you to set your lights to.
The layout is very similar to that of the target matrix. But with extra rows for the timepoints, since some models of lights requires that.

## Exporting the regime

Currently, we can only export regimes compatible with Heliospectra DYNA(TM) lights running the R3.2.2-Release firmware.

```{r}
write.helioSchedule(regime, filename='my_regime.txt', format='json')
```

If you do not use this model of light, you can still write the regime matrix to a csv.
Or you're very welcome to write your own function to format the regime to be compatible with your model of lights!

# How it works

`makeRegime` is the core user-facing function in this package.
But you probably want to know what is happening behind the scenes.
We will summarise below, and full explanation is available at: https://doi.org/10.1101/2025.06.06.658293

When you run `makeRegime`, it carries out 4 steps in the background:

1. Calculate closest intensities
2. Predict the intensities to use to achieve the target irradiance (via a system of linear equations or non-negative least squares)
3. Tidy the intensities (rounding to integer, keep within the range of intensities that the lights can be set to)
4. Format the intensities and timestamps into a human-readable regime matrix

We'll go through each of the steps below.

## Closest intensities

This step searches through the calibration data to find intensities which produce the closest irradiances to your target irradiance.

Using the example target irradiances from above:

```{r}
closest <- LightFitR::internal.closestIntensities(target, calibration[, c(3,5,4,6)])
rownames(closest) <- LightFitR::helio.dyna.leds$name
print(closest)
```

We've got a matrix of intensities for each channel and event!

To convince ourselves that these are indeed the closest, we can compare our target irradiances with the calibration data.
Let's do this for the first event:

```{r}

# Define variables

## Calibration
calib_wavelengths <- unique(calibration$wavelength)
calib_intensities <- unique(calibration$intensity)

## Subset the closest matrix to the first event
closest_first <- closest[,1]
print(closest_first)

## Subset the targets
target_first <- target[,1]
print(target_first)

# Go through each channel of the first event
sapply(1:length(closest_first), function(i){
  
  ## Set relevant variables
  tar <- target_first[i]
  clo <- closest_first[i]
  led <- helio.dyna.leds[i, 'wavelength']
  
  ## Subset calibration data to the LED at the peak wavelengths
  criteria <- calibration$led==led & calibration$wavelength==LightFitR:::internal.closestWavelength(calib_wavelengths, led)
  calib_subset <- calibration[criteria, 3:6]
  
  # Print outputs for user
  print(names(clo))
  print('This is the calibration data')
  print(calib_subset)
  print(paste('The target irradiance is', tar))
  print(paste('The closest intensity is', clo))
  print('---')
  
  return()
})

```

These closest intensities are used in the next step.

## Predict intensities using SLE or NNLS

This step predicts the intensities to use, with a system of linear equations or non-negative least squares (a fancy version of SLE).

First, it uses the closest intensities and calibration data to create a matrix of irradiances for each event. 
Think of this as a heatmap with all of the bleedthrough between all the channels of the LED at the closest intensity.
Again, we'll show this for the first event:

```{r}

# Define variables
peakWavelengths <- LightFitR:::internal.closestWavelength(unique(calibration$wavelength), helio.dyna.leds[-9, 'wavelength'])
firstEvent <- data.frame(led=LightFitR::helio.dyna.leds[-9, 'wavelength'], closest=closest_first[-9], intended=target_first[-9])
rm(closest_first, target_first)

mat <- sapply(1:nrow(firstEvent), function(j){
  criteria <- (calibration$led == firstEvent[j, 'led']) & (calibration$intensity == firstEvent[j, 'closest']) & (calibration$wavelength %in% peakWavelengths) # We want the irradiances (from calibration data) of each LED at the intensity where it is closest to the intended irradiance
  calibration[criteria, 'irradiance']
})

print(mat)
image(mat)

```

The leading diagonal is quite intense (which is what we want!). 
But as you can see, there is some bleedthrough between the channels, indicated by colour outside the leading diagonal.

Next, we make a model either using NNLS or SLE (more on the differences below). 
It solves the equation Ax=b, where A is the irradiance matrix (above), b are the target irradiances and x are unknown coefficients.

For the first event, it looks like this:

```{r}
mod <- nnls::nnls(mat, firstEvent[,'intended'])
print(mod)
```

Finally, we use the coefficients (solved by the model) as well as the closest intensities to calculate the intensities we need to set the lights to:

```{r}
intensities <- mod$x * firstEvent[, 'closest']
print(intensities)
```

Great! We have our intensities! We can put that straight into the lights right?

Well, most lights only accept whole numbers. 
And, although not apparent in this example, sometimes we predict intensities which are impossible for the lights (e.g. above 1000 in the case of Heliospectra DYNAs).
So this brings us onto the next step!

## Tidying

This is exactly as it sounds.
We round our predicted intensities to the nearest integer, cap the predicted intensities to the maximum the lights can achieve (inferred from the calibration data), set any negative predictions to 0.

```{r}
tidied <- LightFitR:::internal.tidyIntensities(intensities, calib_intensities)
print(tidied)
```

Much better!

## Formatting

This final step just takes the resultant matrix and makes it more human readable.
It adds row and column names, as well as the event timepoints.
You can't really show it with just the first event, and you already saw the tidied matrix in the 'get started' example.
So there isn't really any code to show for this step...

# System of linear equations (SLE) vs Non-negative least squares (NNLS)

From our testing, there isn't much of a difference between SLE and NNLS when it comes to predicted intensities (https://doi.org/10.1101/2025.06.06.658293), except for occasional outliers.
So if you're not happy with the intensities predicted by one method, try the other.
The default for the package is NNLS but this was an arbitrary decision on our part.

## SLE

This is the 'simpler' of the two methods.
It solves the equation Ax=b straightforwardly.
However, for us, this means it can predict intensities below 0 (i.e. unachievable by the lights since 0 means the light is off) and intensities above the maximum (again, impossible on the lights).

## NNLS

This uses the Lawson-Hanson method to solve Ax=b in such a way that the predicted intensities are non-negative.
This gets around the below 0 predictions, but can still predict intensities above the maximum.
See the `nnls` package for more info on how NNLS works.
